The role of habitat quality in host metapopulations exposed to environmentally persistent pathogens. Or Identifying when environmentally persistent pathogens transform good habitat into ecological traps

### 2015-06-30 Manuscript/manuscript.tex
Fig 5 to supplement, remove fig 6

To better understand the patch-level mechanisms behind these effects, we further examined within patch infection dynamics for a pathogen in this intermediate range (again with $\nu = 0.2$ and $\delta = 0.3$).  
Because of their lower extinction rate and higher recolonization rate, high quality patches supported more consistent occupancy than low quality patches (Fig~\ref{simvis}).  
However, disease risk, measured by the proportion of time a patch spent occupied by infectious hosts, increased with patch quality, an effect that became more pronounced as pathogen longevity increased (Fig~\ref{pinfection}).
Indeed, as an epidemic progressed, high quality patches increasingly supported infectious occupants, while low quality patches supported infectious hosts only briefly and infrequently (Fig~\ref{simvis}).  
Thus, from the perspective of a migrant individual, colonizing high quality habitat posed increased disease risk, with the preferential selection of high quality habitat (i.e. $\xi_{im} = 0.5$) increasing this disease risk further (Fig~\ref{pinfection}, a-c vs. d-f).
Moreover, preferential selection of high quality habitat had dramatic consequences for the stability of the host metapopulation when the pathogen is long-lived in the environment, increasing extinction risk substantially compared to random habitat selection (i.e. $\xi_{em} = 0$, Fig~\ref{pext}).

@@

\begin{figure}
%\includegraphics[scale=0.4]{figure/figure_5.pdf}
\centering
\caption{Boxplots showing the probability of infection (i.e. proportion of time occupied by infectious hosts) as a function of patch quality.  Panel columns show low (0.3, a, d), medium (1.0, b, e), and high (3.2, c, f) pathogen longevities, while rows show the effect of preferential selection of high quality habitat (a-c, $\xi_{im} = 0$,  no preference; d-f, $\xi_{im} = 0.5$, preference for high quality). The bottom row demonstrates the existence of an ecological trap on high quality habitat resulting from the fitness costs of increased disease risk. (PCC: delete sentence goes in the main text, but it also isn't clear to me how you derive this conclusion from the figure) Black points show the median over 100 replicate simulations, while vertical lines show the outer-quartile range, and gray points indicate outliers. Disease induced mortality and probability of direct infection fixed at $\nu = 0.2$ and $\delta = 0.3$. PCC: you can scale y to be the same across all plots, which would allow you to drop your y axes on col 2 \& 3, same goes for dropping the x axis on row 1 and squishing it together. Again, I would lable on the plot itself how longevity runs acroll the columns and rows indicate preference. If your main intent of the figure is to compare zero vs. high preference, then you should have those 2 cases on the same plot (maybe as two lines, dropping the boxplots) rather than making the reader make this connection across plots. The information in the boxplot seems to distract, rather than contribute, to your message here. }
\label{pinfection}
\end{figure}

\begin{figure}
%\includegraphics[scale=1]{figure/figure_6.pdf}
\centering
\caption{Probability of host extinction as a function of pathogen environmental longevity and habitat preference, with black showing no habitat preference ($\xi_{im} = 0$) and red showing preference for high quality habitat ($\xi_{im} = 0.5$).  Disease induced mortality and probability of direct infection fixed at $\nu = 0.2$ and $\delta = 0.3$.  The combination of preference for high quality habitat and an environmentally long-lived pathogen serve to substantially increase extinction risk.  (PCC: seems like you could say this in text w/o using a figure since you have only 6 points here and only 2 of them are different....i.e. The probability of host extinction was ten times higher when there was a preference for high quality patches and longer environmental persistence (0.3 vs. 0.03))}
\label{pext}
\end{figure}
\clearpage

>>>>

To better understand the patch-level mechanisms behind these effects, we further examined within patch infection dynamics for a pathogen in this intermediate range (again with $\nu = 0.2$ and $\delta = 0.3$).  
Because of their lower extinction rate and higher recolonization rate, high quality patches supported more consistent occupancy than low quality patches (Fig~\ref{simvis}, Supplement).  
However, because of this stability, disease risk, measured by the proportion of time a patch spent occupied by infectious hosts, increased with patch quality (Supplement).
Indeed, as an epidemic progressed, high quality patches increasingly supported infectious occupants, while low quality patches supported infectious hosts only briefly and infrequently (Fig~\ref{simvis}).  
Thus, from the perspective of a migrant individual, colonizing high quality habitat posed increased disease risk, with the preferential selection of high quality habitat (i.e. $\xi_{im} = 0.5$) increasing this disease risk further (Supplement).
Moreover, assuming preferential selection of high quality habitat had dramatic consequences for the stability of the host metapopulation when the pathogen is long-lived in the environment, increasing the probability of extinction substantially compared to random habitat selection (0.29 vs. 0.03).

@@
